---
title: "Part 5 - LDSC with external trait"
author: "Wangjingyi Liao"
date: "2024-02-14"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

1. LDSC and jack knife corrected genetic correlation comparision - script adapted from Joni Coleman.

check the ./LDSC_external_trait_pipeline folder

useage:
  1. munge summary statistics for external traits and save them to the munged_external_binary and munged_External_count folder, binary and continuous traits will be analysed separately. 
  2. munged original summary statistics uncorrected for p and non-p summary statistics saved to the p_sumstats and non-p_sumstats folder
  3. make sure the ldsc is actual kept in the ldsc folder: e.g. list ./ldsc/ldsc 
  4. cd to the ldsc directory
  5. open the WrapperScript.sh file and edit the working directories and folder directories 
  6. within a ldsc environment (e.g. conda activate ldsc), submit the sh file using: sh WrapperScript.sh
  7. individual outputs will be saved in to the jf_results folder and LDSC log files will be saved into Logs folder




2. Making Figure 3-4 - visualization of the genetic correlation with external trait


```{r}
library(tidyverse)
library(reshape2)

```

```{r}
rm (list = ls())

# Load necessary libraries
library(data.table)
library(dplyr)

# Read data from the CSV file into the data frame 'external_rgs'
setwd('./robustness_check/ldsc_nonp_externaltrait_residuals/ldsc/')
# Read the file using fread and skip lines with empty cells
external_rgs <- fread('./p_nonp_allexternal_traits_ldsc.txt', fill = T)

# Remove rows with NA
rgs <- na.omit(external_rgs)

```

```{r}
head(rgs)
class(rgs$ExternalTrait)
rgs$ext_tr<- as.factor(rgs$ExternalTrait)
levels(rgs$ext_tr)

# #use the code below to rename the levels to look consistent and interpretable
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Age_at_smoking_initiation"] <- "Age at smoking initiation"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Menopause"] <- "Age at menopause"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="birth_weight" ] <- "Birth weight"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Childhood_IQ" ] <- "Childhood IQ"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="cigday" ] <- "Cigarette per day"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="drinkweek" ] <- "Drink per week"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="EA4" ] <- "Education attainment"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="executivefunctioning" ] <- "Executive function"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="household_income" ] <- "Household income"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="intelligence2"] <- "Cognitive performance"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="lonelines"] <- "Lonelines"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="longevity"] <- "Longevity"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Metabolic_Syndrome"] <- "Metabolic syndrome"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="neighbourhood_deprivation"] <- "Neighbourhood deprivation"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="no_sexual_partners"] <- "Number of sexual partners"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Risk_beh_composite"] <- "Risk behavior composite"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="risk_tolerance"] <- "Risk tolerance"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="selfrated_health"] <- "Self-rated poor health"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="sleepduration"] <- "Sleep duration"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Speeding_propensity"] <- "Speeding propensity"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="subjective_wellbeing"] <- "Subjective well-being"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Tiredness"] <- "Tiredness"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Waist_to_hip_ratio"] <- "Waist to hip ratio"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="environmentalstress"] <- "Environmental stress"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="noncog"] <- "Noncognitive skills"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="internalizing"] <- "Internalizing problem"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="externalizing"] <- "Externalizing problem"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="neurodegenerative"] <- "Neurodegenerative disorders"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="backpain"] <- "Back pain"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="insomnia"] <- "Insomnia"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="CUD"] <- "Cannabis use disorder"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="ever_smoker"] <- "Ever smoker"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="Smoking_cessation"] <- "Smoking cessationn"
levels(rgs$ext_tr)[levels(rgs$ext_tr)=="chronotype"] <- "Chronotype"

# FDR correction for the correction comparision 
# FDR correction for the correction comparision 
rgs$fdr_jf_p <- p.adjust(rgs$jf_p, method = "fdr") #  146 significant results

# Subset rgs with fdr_jf_p < 0.05
rgs_sig_fdr <- rgs[rgs$fdr_jf_p < 0.05, ]

order <- as.character(sort(levels(rgs$ext_tr)))

rgs$ext_tr <- factor(rgs$ext_tr, levels = order)

## Define categories (loosely based on the GWAS-by-subtraction but changes made in order to fit sum stats availability)
demographic<- c("Household income", 
                "Neighbourhood deprivation",
                "Longevity")
health<-c("Risk behavior composite",
          "Risk tolerance",
          "Speeding propensity",
          "Number of sexual partners" ,  
          "Ever smoker",
          "Drinks per week", 
          "Sleep duration" ,
          "Number of sexual partners",
          "Age at smoking initiation", 
          "Chronotype", 
          "Cigarette per day",
          "Diabetes T2",
          "Drink per week" ,
          "Metabolic syndrome",
          "Self-rated poor health",
          "Neurodegenerative disorders",
          "Back pain",  
          "Age at menopause", 
          "Smoking cessationn")
psychological<- c("Childhood IQ",
                  "Noncognitive skills",
                  "Education attainment",
                  "Executive function",
                  "Neuroticism",
                  "Environmental stress",
                  "Subjective well-being", 
                  "Tiredness", 
                  "Cognitive performance",
                  "Lonelines",
                  "Internalizing problem",
                  "Externalizing problem",
                  "Insomnia", 
                  "Cannabis use disorder")
antropometric<-c("Birth weight", 
                 "Waist to hip ratio"
                 )



# add a column with the category name 
rgs$category <- NULL
rgs$category[rgs$ext_tr %in% health] <- "Health related"
rgs$category[rgs$ext_tr %in% psychological] <- "Psychological"
rgs$category[rgs$ext_tr %in% demographic] <- "Demographic"
rgs$category[rgs$ext_tr %in% antropometric] <- "Anthropometric"

## To fix the order of the groups and traits 
rgs$category_f <- factor(rgs$category, levels=c("Demographic", "Anthropometric","Health related","Psychological")) 

rgs$disorder <- rgs$Trait

datl <- rgs
```




```{r}
###Health-related traits###


health <- datl[which(datl$category_f%in% c("Health related")), ] 

#nudge <- position_nudge(x = -1, y = 0)
nudge1 <- position_nudge(x = .1, y = 0)
nudge2 <- position_nudge(x = -.1, y = 0)

rghealth<- ggplot(health, aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)))) + 
  scale_colour_manual(name="Data", values=c("0" = "#060b5c","1" = "#66CDAA", "2" = "red"), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) + #need trick with both values and labels and number to order the legend.....
  scale_shape_manual(name="Data", values=c("0" =16,"1" =16, "2" =8), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) +
  geom_point(aes(y=p_cor, x=reorder(ext_tr, desc(ext_tr)),color="0", shape="0"), size=3, position = nudge1) +
  geom_errorbar(aes(ymin=p_cor-1.96*p_cor_se, ymax=p_cor+1.96*p_cor_se,color="0"), width=.1, show.legend=FALSE, position = nudge1) +
  geom_point(aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)),color="1", shape="1"), size=3, position = nudge2)+
  geom_errorbar(aes(ymin=nonp_cor-1.96*nonp_cor_se, ymax=nonp_cor+1.96*nonp_cor_se,color="1"), width=.1, show.legend=FALSE, position = nudge2) +
  geom_point(data = health[health$fdr_jf_p < .05, ],size=2, aes(y = -.90, x = reorder(ext_tr, desc(ext_tr)), color="2",shape = '2')) + # star to indicate the significance of the difference 
  ylim(-1, 1) +
  geom_hline(yintercept = 0) +
  theme_light(base_size=14) +
  theme(axis.title.x=element_blank(),
        #axis.title.y=element_blank(),
        legend.title=element_blank()) +
  ylab("Genetic Correlation") +
  xlab("")+
  theme(legend.position="bottom") +
  facet_grid(. ~ category_f, scales = "free", space = "free_y") +
  coord_flip() 
  #scale_x_discrete(limits = rev)# flip y and x 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # adjust the x axis text 45 degree


rghealth1 = rghealth+facet_wrap(vars(disorder), ncol = 6)+
  theme(strip.text = element_text(size=10, colour = 'black', face = "bold"),
      strip.background = element_rect(colour="gray", fill="white"))


plot(rghealth1)

ggsave(rghealth1, file="./robustness_check/ldsc_nonp_externaltrait_residuals/figures_ldsc/p_nonp_rgs_health.tiff", width=13, height=11)

```

###Psychological traits###
```{r}
psychol <- datl[which(datl$category_f%in% c("Psychological")), ] 
dim(psychol)
psychol<-psychol[!(psychol$ExternalTrait=="Cog_demange"),]
dim(psychol)
#nudge <- position_nudge(x = -1, y = 0)
nudge1 <- position_nudge(x = .1, y = 0)
nudge2 <- position_nudge(x = -.1, y = 0)

rgpsych<- ggplot(psychol, aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)))) + 
  scale_colour_manual(name="Data", values=c("0" = "#060b5c","1" = "#66CDAA", "2" = "red"), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) + #need trick with both values and labels and number to order the legend.....
  scale_shape_manual(name="Data", values=c("0" =16,"1" =16, "2" =8), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) +
  geom_point(aes(y=p_cor, x=reorder(ext_tr, desc(ext_tr)),color="0", shape="0"), size=3, position = nudge1) +
  geom_errorbar(aes(ymin=p_cor-1.96*p_cor_se, ymax=p_cor+1.96*p_cor_se,color="0"), width=.1, show.legend=FALSE, position = nudge1) +
  geom_point(aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)),color="1", shape="1"), size=3, position = nudge2)+
  geom_errorbar(aes(ymin=nonp_cor-1.96*nonp_cor_se, ymax=nonp_cor+1.96*nonp_cor_se,color="1"), width=.1, show.legend=FALSE, position = nudge2) +
  geom_point(data = psychol[psychol$fdr_jf_p < .05, ],size=2, aes(y = -.90, x = reorder(ext_tr, desc(ext_tr)), color="2",shape = '2')) + # star to indicate the significance of the difference 
  ylim(-1, 1.2) +
  geom_hline(yintercept = 0) +
  theme_light(base_size=14) +
  theme(axis.title.x=element_blank(),
        #axis.title.y=element_blank(),
        legend.title=element_blank()) +
  ylab("Genetic Correlation") +
  xlab("")+
  theme(legend.position="bottom") +
  facet_grid(. ~ category_f, scales = "free", space = "free_y") +
  coord_flip() # flip y and x 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # adjust the x axis text 45 degree


rgpsych1 = rgpsych+facet_wrap(vars(disorder), ncol = 6)+
  theme(strip.text = element_text(size=10, colour = 'black', face = "bold"),
      strip.background = element_rect(colour="gray", fill="white"))


plot(rgpsych1)

ggsave(rgpsych1, file="./robustness_check/ldsc_nonp_externaltrait_residuals/figures_ldsc/p_nonp_rgs_psychological.tiff", width=13, height=11)

```

```{r}
demographic <- datl[which(datl$category_f%in% c("Demographic")), ] 
demographic <- demographic %>% 
  arrange(ext_tr)

#nudge <- position_nudge(x = -1, y = 0)
nudge1 <- position_nudge(x = .1, y = 0)
nudge2 <- position_nudge(x = -.1, y = 0)

fig_demographic <- ggplot(demographic, aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)))) + 
  scale_colour_manual(name="Data", values=c("0" = "#060b5c","1" = "#66CDAA", "2" = "red"), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) + #need trick with both values and labels and number to order the legend.....
  scale_shape_manual(name="Data", values=c("0" =16,"1" =16, "2" =8), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) +
  geom_point(aes(y=p_cor, x=reorder(ext_tr, desc(ext_tr)),color="0", shape="0"), size=3, position = nudge1) +
  geom_errorbar(aes(ymin=p_cor-1.96*p_cor_se, ymax=p_cor+1.96*p_cor_se,color="0"), width=.1, show.legend=FALSE, position = nudge1) +
  geom_point(aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)),color="1", shape="1"), size=3, position = nudge2)+
  geom_errorbar(aes(ymin=nonp_cor-1.96*nonp_cor_se, ymax=nonp_cor+1.96*nonp_cor_se,color="1"), width=.1, show.legend=FALSE, position = nudge2) +
  geom_point(data = demographic[demographic$fdr_jf_p < .05, ],size=2, aes(y = -.90, x = reorder(ext_tr, desc(ext_tr)), color="2",shape = "2")) + # star to indicate the significance of the difference 
  ylim(-1, 1) +
  geom_hline(yintercept = 0) +
  theme_light(base_size=14) +
  theme(axis.title.x=element_blank(),
        #axis.title.y=element_blank(),
        legend.title=element_blank()) +
  ylab("Genetic Correlation") +
  xlab("")+
  theme(legend.position="bottom") +
  facet_grid(. ~ category_f, scales = "free", space = "free_y") +
  coord_flip() 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # adjust the x axis text 45 degree


fig_demographic1 = fig_demographic+facet_wrap(vars(disorder), ncol = 6)+
  theme(strip.text = element_text(size=10, colour = 'black', face = "bold"),
      strip.background = element_rect(colour="gray", fill="white"))


plot(fig_demographic1)

ggsave(fig_demographic1, file="./robustness_check/ldsc_nonp_externaltrait_residuals/figures_ldsc/p_nonp_rgs_demographic.tiff", width=13, height=9)

```


```{r}
anthro <- datl[which(datl$category_f%in% c("Anthropometric")), ] 

#nudge <- position_nudge(x = -1, y = 0)
nudge1 <- position_nudge(x = .2, y = 0)
nudge2 <- position_nudge(x = -.2, y = 0)

rganthro<- ggplot(anthro, aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)))) + 
  scale_colour_manual(name="Data", values=c("0" = "#060b5c","1" = "#66CDAA", "2" = "red"), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) + #need trick with both values and labels and number to order the legend.....
  scale_shape_manual(name="Data", values=c("0" =16,"1" =16, "2" =8), labels = c("Uncorrected for p", "Non-p", "FDR-corrected \nsignificant difference \nbetween pairs of correlations")) +
  geom_point(aes(y=p_cor, x=reorder(ext_tr, desc(ext_tr)),color="0", shape="0"), size=3, position = nudge1) +
  geom_errorbar(aes(ymin=p_cor-1.96*p_cor_se, ymax=p_cor+1.96*p_cor_se,color="0"), width=.1, show.legend=FALSE, position = nudge1) +
  geom_point(aes(y=nonp_cor, x=reorder(ext_tr, desc(ext_tr)),color="1", shape="1"), size=3, position = nudge2)+
  geom_errorbar(aes(ymin=nonp_cor-1.96*nonp_cor_se, ymax=nonp_cor+1.96*nonp_cor_se,color="1"), width=.1, show.legend=FALSE, position = nudge2) +
  geom_point(data = anthro[anthro$fdr_jf_p < .05, ],size=2, aes(y = -.90, x = reorder(ext_tr, desc(ext_tr)), color="2",shape = '2')) + # star to indicate the significance of the difference 
  ylim(-1, 1) +
  geom_hline(yintercept = 0) +
  theme_light(base_size=14) +
  theme(axis.title.x=element_blank(),
        #axis.title.y=element_blank(),
        legend.title=element_blank()) +
  ylab("Genetic Correlation") +
  xlab("")+
  theme(legend.position="bottom") +
  facet_grid(. ~ category_f, scales = "free", space = "free_y") +
  coord_flip() 
  #guides(size="none", shape="none")# flip y and x 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # adjust the x axis text 45 degree


rganthro1 = rganthro+facet_wrap(vars(disorder), ncol = 6)+
  theme(strip.text = element_text(size=10, colour = 'black', face = "bold"),
      strip.background = element_rect(colour="gray", fill="white"))


plot(rganthro1)


ggsave(rganthro1, file="./robustness_check/ldsc_nonp_externaltrait_residuals/figures_ldsc/p_nonp_rgs_anthropometric.tiff", width=11, height=7)

```




